---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

Load necessary libraries:

```{r message=FALSE}
library(dplyr)
library(lattice)
```

## Loading and preprocessing the data

1. Extract the data if doesnt extracted and load:

```{r}
if(!file.exists("activity.csv")){
    unzip ("activity.zip", exdir = "./")
}
data<-read.csv("activity.csv")
```

2. Convert date column to Date Class:

```{r}
data$date<-as.Date(data$date)
```

## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day

```{r}
stepsPerDay<-with(data, tapply(steps, date, sum))
```

2. Make a histogram of the total number of steps taken each day

```{r}
hist(stepsPerDay, main="Histogram of Total Number of Steps Each Day", xlab="Number of Steps")
```

3. Mean and median of the total number of steps taken per day

```{r}
summary(stepsPerDay)[c("Mean","Median")]
```

## What is the average daily activity pattern?

1. Make a time series plot (i.e. type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
stepsPer5Minutes<-with(data, tapply(steps, interval, mean, na.rm=TRUE))
stepsPer5Minutes<-as.data.frame.table(stepsPer5Minutes)
names(stepsPer5Minutes)<-c("Minutes", "Mean")
plot(stepsPer5Minutes$Minutes, stepsPer5Minutes$Mean, type="l", main="Average of steps with in intervals", xlab="Minutes", ylab="Average number of steps")
lines(stepsPer5Minutes$Minutes, stepsPer5Minutes$Mean, type="l", col="red")
```

2. The 5-minute interval that, on average, contains the maximum number of steps

```{r}
as.numeric(as.character(stepsPer5Minutes$Minutes))[which.max(stepsPer5Minutes$Mean)]
```

## Imputing missing values

1. Total number of missing values in the dataset

```{r}
sum(is.na(data))
```

2. Missing values will be filled with average number of steps taken each 5-minute interval.

Get means of all 5 minutes interval which has missing value

```{r}
missingValues<-stepsPer5Minutes$Mean[stepsPer5Minutes$Minutes %in% data$interval[is.na(data$steps)]]
```

3. Create a new dataset with missing data filled

```{r}
filledData<-data.frame(data)
filledData$steps[is.na(filledData$steps)]<-missingValues
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}
stepsPerDay<-with(filledData, tapply(steps, date, sum))
hist(stepsPerDay, main="Histogram of Total Number of Steps Each Day", xlab="Number of Steps")
summary(stepsPerDay)[c("Mean","Median")]
```

Mean is not changed and median is slightly increased. 

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
filledData<-mutate(filledData, day_type = 
    factor(case_when(weekdays(data$date)==c("Saturday", "Sunday") ~ "weekend", TRUE ~ "weekday")))
```

2. Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
filledData_interval_means <- aggregate(steps ~ interval + day_type, filledData, mean)
xyplot(steps ~ interval | day_type,
       data = filledData_interval_means,
       type = "l",
       xlab = "Interval",
       ylab ="Number of steps",
       layout=c(1,2))
```


